---
title: "Dplyr Tutorial"
author: "Rick"
date: "27/01/2021"
output: 
  html_document:
  code_folding: "show"
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r load, Libraries}
library(tidyverse)
library(knitr)
library(nycflights13)

```

In this tutorial we will explore the Dplyr package of the Tidyverse. In this tutorial we will examine the following functions, along with the associated cheat sheet.

1) filter()
2) arrange()
3) select()
4) rename()
5) mutate() and extensions
6) group_by
7) summarise()
8) left_join()

Please niote that this tutorial was adapted from chapters 5 and 13 of the book 'R for Data Science' by Hadley Wickham and Garrett Grolemund. This book can be found at https://r4ds.had.co.nz/ The tutorial was originally delivered by RichardonData and the tuorial can be found at https://www.youtube.com/watch?v=6Z48qcrTfYU

A good cheat sheet for dplyr functions canbe found at https://rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf

```{r load data}
flights <- flights
airlines <- airlines
```

###Filter###

We can use the 'filter()' function to filter rows of our datasets based on specific conditions. 

Note that we use the 'pipe' operator '%>%' here. This is a fairly straightforward syntax associated with the tidyverse. You can essentially read this like the word 'then'. In the example below take the dataset called 'flgiths' and filter it. Multiple operations can be piped together as we will see later. 

Also note that the prefix 'dplyr::' will be used in front of some of these functions to avoid clashes with similar named functions in other packages. 

```{r example of "filter"}
flights %>%
  dplyr::filter(month == 1 | month == 2, day == 1)
```

Next, we will pipe multiple operations together. 

```{r longer example of "filter"}
flightsFiltered <- flights %>%
  dplyr::filter(month %in% c(11, 12)) %>%
  dplyr::filter(dep_time >= 700) %>%
  dplyr::filter(carrier != 'UA')
flightsFiltered
```

<br>

###Arrange###

The 'arrange()' function can be used to sort a data frame. Multiple variables can be taken in this function, and the desc() helper can be used to specifcy decending rather than ascending order. 

```{r example of "arrange"}
flightsArranged <- flightsFiltered %>%
  arrange(year, month, day, desc(dep_delay))
flightsArranged
```

<br>

###Select and Rename###

The 'select()' function can be used to select variables (i.e. columns) of the dataset. A '-' sign can be used to remove columns. 
The 'rename()' function can be used to change the name of variables in a daatset. 

```{r examples of "select" and "rename"}
flightsSelected <- flightsArranged %>%
  dplyr::select(-(hour:time_hour)) %>%
  dplyr::select(flight, year, month, day, dep_time, dep_delay, arr_time, arr_delay, carrier, origin, dest, air_time, distance) %>%
  dplyr::rename("airtime" = "air_time", "destination" = "dest")
flightsSelected
```

<br>

###Mutate###

The "mutate()" function can be used to create new variables.

```{r example of "mutate"}
flightsMutated <- flightsSelected %>%
  mutate(gain = dep_delay - arr_delay,
         hours = airtime / 60,
         gain_per_hour = gain/hours)
flightsMutated
```

There are various extensions of the mutate function. The old helper functions 'mutate_at()' and 'mutate_all()', which can be used to modify multiple variables at a time. The arguements for these functions were '.vars' and '.funs'. Recently is the '.across()' function which will take in arguements .cols and .fns. This is the recommended way of extending the 'mutate()' function. 

```{r extensions of "mutate"}
flight2A <- flightsMutated %>%
  mutate_at(.vars = c("year", "month", "day"), .funs = as.character)
flights2B <- flightsMutated %>%
  mutate(across(.cols = c("year", "month", "day"), .fns = as.character))


