---
title: "Dplyr Tutorial"
author: "Rick"
date: "27/01/2021"
output: 
  html_document:
  code_folding: "show"
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r load, Libraries}
library(tidyverse)
library(knitr)
library(nycflights13)

```

In this tutorial we will explore the Dplyr package of the Tidyverse. In this tutorial we will examine the following functions, along with the associated cheat sheet.

1) filter()
2) arrange()
3) select()
4) rename()
5) mutate() and extensions
6) group_by
7) summarise()
8) left_join()

Please niote that this tutorial was adapted from chapters 5 and 13 of the book 'R for Data Science' by Hadley Wickham and Garrett Grolemund. This book can be found at https://r4ds.had.co.nz/ The tutorial was originally delivered by RichardonData and the tuorial can be found at https://www.youtube.com/watch?v=6Z48qcrTfYU

A good cheat sheet for dplyr functions canbe found at https://rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf

The 


